---
title: "Problem Set 4"
author: "Etienne Danis"
date: "3/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

# Submission

1. Fork this repository to your own account
1. Make changes to the `README.Rmd` file (including the author field above).
1. Knit the file to HTML format
1. Publish HTML on Rpubs (you will need to sign up for the free account) and *add the link below*.
1. Submit a PR with your answers.

*Link to Rpubs document:* XXX

# Overview

Take the [datacamp course on joining data](https://www.datacamp.com/courses/joining-data-in-r-with-dplyr) to learn the join verbs.

You will analyze some data in the [`nihexporter`](https://github.com/jayhesselberth/nihexporter) package to answer some questions. The questions will cover methods in `dplyr` including two-table verbs.

First you need to install the `nihexporter` package (and `devtools`, if you don't have it already).

```r
# install.packages('devtools')
devtools::install_github("jayhesselberth/nihexporter")
```

Read the [Overview vignette](http://jayhesselberth.github.io/nihexporter/articles/nihexporter.html) to learn the structure of the package. Take a look at two tables from the package:

```{r overview}
library(nihexporter)
library(tidyverse)
library(broom)

projects
project_io
```

# Problems

## Problem 1

The `projects` and `project_io` tables have the `r intersect(names(projects), names(project_io))` column in common. Use this column to perform a `left_join` and identify the most expensive project for each institute. Project institutes are in the `projects` table, whereas total project costs are available in the `project_io` table. Note you will have to use the `arrange` and `slice` verbs from `dplyr`.

```{r problem 4.1}
projects %>% select(project.num, institute, fy.cost) %>% 
  left_join(project_io) %>% 
  select (project.num, institute, fy.cost, total.cost) %>% 
  arrange(desc(total.cost)) %>% 
  group_by(institute) %>% 
  slice(1:1) %>% 
  arrange(desc(total.cost)) -> most_costly_per_institute
 most_costly_per_institute
```

## Problem 2

Create a plot that describes the change in output (number of publications) for the most expensive grant from the above list. Note that you cannot use `n.pubs` for this, as that is an aggregate count over time. You need a tbl of following structure, which you can then use to calculate the number of publications each year (hint: use `n()` in a `summarize` statement:

```{r problem 4.2}
most_costly_per_institute %>% head(1) %>% 
  left_join(publinks) %>%
  left_join(publications) %>%
  select(project.num, institute, total.cost, pub.year, pmid) %>%
  group_by(project.num, institute, total.cost, pub.year) %>% summarize(total.pubs= n()) %>% 
  na.omit() -> most_costly_publications_over_time

ggplot(most_costly_publications_over_time, aes(pub.year, total.pubs)) + geom_point()
```



## Problem 3

Use the `projects` and `clinical_studies` table to calculate how many clinical studies have been started by projects from each institute.

```{r problem 4.3}
projects %>% select(project.num, institute) %>% 
  left_join(clinical_studies) %>%
  select(project.num, institute, trial.id) %>%
  group_by(project.num, institute) %>% summarize(total.clin.studies=n()) %>%
  na.omit() %>%  arrange(desc(total.clin.studies))
  
```

## Problem 4

Calculate correlations between `total.cost` and `n.pubs` for each institute. You will need to `join` the `projects` and `project_io` tables to link `institute` with `total.cost` and `n.pubs`. You will then need to use `group_by` and `do` from `dplyr` and `tidy` from `broom` to calculate correlations.


```{r problem 4.4.1}
# Solution based on Jay's recommendations 
projects %>% left_join(project_io) %>% select(institute, total.cost, n.pubs) %>% na.omit() %>% 
  group_by(institute) %>% summarize(institute.cost=sum(total.cost), institute.pubs = sum(n.pubs)) %>% 
  do(tidy(cor(.$institute.cost, .$institute.pubs))) -> coeff_correlation_cost_pubs
  coeff_correlation_cost_pubs
```

### The coefficient of correlation between the total funding received by an institute and the total number of publications published by an institute is `r coeff_correlation_cost_pubs`!!!

```{r problem 4.4.2}
# Alternative code described by one of the students in the class
projects %>% left_join(project_io) %>% select(institute, total.cost, n.pubs) %>% na.omit() %>% 
  group_by(institute) %>% summarize(institute.cost=sum(total.cost), institute.pubs = sum(n.pubs)) %>% summarize(cor(institute.cost,institute.pubs))  
# Conclusion: Using summarize() twice in a row can be a great substitution to use a combination of the functions do() and tidy()
```

```{r problem 4.4.3 just by curiosity}
# The correlation is so good that I had to see it on a graph
library(cowplot)
projects %>% left_join(project_io) %>% select(institute, total.cost, n.pubs) %>% na.omit() %>% 
  group_by(institute) %>% summarize(institute.cost=sum(total.cost), institute.pubs = sum(n.pubs)) -> data_for_graph_cost_npubs_by_institute
ggplot(data_for_graph_cost_npubs_by_institute, aes(institute.cost, institute.pubs)) + geom_point() + theme_cowplot() + 
  ylab('Total number of publications per institute') + 
  xlab('Total funding received by institute') + 
  ggtitle('Total number of publications per institute 
  reported by total funding received by institute')
# Conclusion: The more funding an institute receives, the more publications the institute produces.
```



## Problem 5 (extra credit)

Correlate [`rcr`](https://icite.od.nih.gov/) values from the `publications` table to one of:

- total cost (overall for NIH, or by institute)
- budget mechanism (e.g., RFA, PAR, etc.)
- type of grant (e.g. R01, P01, R21, etc.). 

```{r extra_credit}
```

## Problem 6 (extra extra credit)

Use [`gganimate`](https://github.com/dgrtwo/gganimate) to illustrate the change in cost of R01s for a few institutes over time.

```{r extra_extra_credit}
```
